---
title: "7-Activity-Joins-Pivots_dataRetrieval"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the tidyverse, dataRetrieval, and patchwork packages.

```{r}
library(tidyverse)
library(dataRetrieval)
library(patchwork)
```

### 1. Using readNWISqw(), read all the chloride (00940) data for the New River at Radford (03171000).

### Use the head() function to print the beginning of the output from readNWISqw.

```{r}
#Chloride and Nitrate
param <- "00940"
site <- "03171000"

newriverWQ <- readNWISqw(siteNumber = site, 
                       parameterCd = param) 

head(newriverWQ)
```

### 2. Using the readNWISdv (daily values) function, download discharge (00060), temperature (00003), and specific conductivity (00095) for the New River at Radford from 2007 to 2009 (regular year).

### Use renameNWIScolumns() to rename the output of the download.

### Use head() to show the beginning of the results of your download.

```{r}
start <- "2007-01-01"
end <- "2009-01-01"
params <- c("00060", "00003","00095")

newphys <- readNWISdv(siteNumber = site, 
                       parameterCd = params, 
                       startDate = start, 
                       endDate = end) %>% 
            renameNWISColumns()

head(newphys)
```

### 3. Do a left join on newphys and newriver to add the chloride data to the daily discharge, temp, and conductivity data. hint: you will join on the date.

### Preview your data below the chunk using head()

```{r}
newwqp <- left_join(newphys, newriverWQ, by = c("Date" = "sample_dt"))
```

### 4. Create a line plot of Date (x) and Flow (y). Create a scatter plot of Date (x) and chloride concentration (y). Put the graphs on top of each other using the patchwork library.

```{r}
QP <- ggplot(newwqp, aes(Date, Flow))+
  geom_line()

WQP <- ggplot(newwqp, aes(Date, result_va))+
  geom_point()

QP/WQP

```

### 5. Create a scatter plot of Specific Conductivity (y) and Chloride (x). Challenge: what could you do to get rid of the warning this plot generates about NAs.

```{r}
ggplot(newwqp, aes(x = SpecCond, y = result_va))+
  geom_point()
```

### 6. Read in the GG chem subset data and plot Mg_E1 (x) vs Ca_E1 (y) as points.

```{r}

subset <- read_csv("GG_chem_subest.csv")

ggplot(subset, aes(x = Mg_E1, y = Ca_E1 ))+
  geom_point()
```

### 7. We want to look at concentrations of each element in the \#6 dataset along the stream (Distance), which is difficult in the current format. Pivot the data  into a long format, the data from Ca, Mg, and Na \_E1 columns should be pivoted.

### Make line plots of each element where y is the concentration and x is distance. Use facet_wrap() to create a separate plot for each element and use the "scales" argument of facet_wrap to allow each plot to have different y limits.

```{r}
wqlong <- subset %>% pivot_longer(cols = c(Ca_E1, Mg_E1, Na_E1))

ggplot(data = wqlong, aes(x = Distance, y = value, color = name))+
  geom_line()+
  facet_wrap(facets = "name", nrow = 3, scales = "free")
```
